---
title: "Reproducible Research Project 1"
author: "Guoliang Ding"
date: "July 15, 2015"
output: html_document
---

## Loading and preprocessing the data

```{r}
library(lubridate)
setwd("C:/Users/kilading/Documents/RLearning/ReproducibleResearch")
activity <- read.csv(file="activity.csv", stringsAsFactors = FALSE)
activity$date <- ymd(activity$date)
activity <- transform(activity, interval = factor(interval))
head(activity)
```


## What is the total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
library(dplyr)
date_activity <- group_by(activity, date)
summarize(date_activity, sum = sum(steps, na.rm = TRUE))
```

2. Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
sum_activity <- summarize(date_activity, sum = sum(steps, na.rm = TRUE))
plot1 <- ggplot(sum_activity, aes(sum)) + geom_histogram(binwidth = 2500)
plot1
```

3. Caluclate the mean and median of the total number taken per day
```{r}
summarize(date_activity, mean = mean(steps, na.rm = TRUE), median = median(steps, na.rm = TRUE))
```

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
interval_activity <- group_by(activity, interval)
sum_interval <- summarize(interval_activity, interval_mean = mean(steps, na.rm = TRUE))
plot2 <- ggplot(sum_interval, aes(as.numeric(interval), interval_mean)) + geom_line() + labs(x = "Interval", y = "Average acorss all day")
plot2
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_position <- which.max(sum_interval$interval_mean)
sum_interval$interval[max_position]
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
good <- sum(complete.cases(activity))
nrow(activity) - good
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in

The statergy applied here is to use the mean for that 5-minute interval

```{r}
for (idx in seq(1, nrow(activity))) {
    
    if (is.na(activity[idx, 1])) {
        
        for (idx2 in seq(1, nrow(sum_interval))) {
            
            if (as.numeric(activity[idx, 3]) == as.numeric(sum_interval[idx2, 1])) {
                
                activity[idx, 1] = sum_interval[idx2, 2]
                
                break
                
                
            }
        }
        
        
    }
    
}

head(activity)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
# make the histrgram
date_activity <- group_by(activity, date)
sum_activity <- summarize(date_activity, sum = sum(steps, na.rm = TRUE))
plot1 <- ggplot(sum_activity, aes(sum)) + geom_histogram(binwidth = 2500)
plot1

# Calculate the mean and median per day
summarize(date_activity, mean = mean(steps, na.rm = TRUE), median = median(steps, na.rm = TRUE))

```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
determine_day_type <- function(day_of_week) {
    
    day_type <- c()
    
    for (a_day in day_of_week) {
        
        if (a_day == 1 | a_day == 7) { day_type <- c(day_type, "weekend")}
        
        else {day_type <- c(day_type, "weekday")}
    }
    
    return(day_type)
    
}

activity <- mutate(activity, day_of_week = wday(date), day_type = determine_day_type(day_of_week))

head(activity)

activity <- transform(activity, day_type = factor(day_type))

activity <- group_by(activity, day_type, interval)

summarized <- summarize(activity, interval_mean = mean(steps, na.rm = TRUE))

plot2 <- ggplot(summarized, aes(as.numeric(interval), interval_mean)) + geom_line() + labs(x = "Interval", y = "Average acorss all day") +  facet_grid( day_type~. )

plot2
```